Interpretable Neural Networks using EAGGA

### 2025-03-14 paper/main.tex
some changes to bullets, started finalising interpretability wording

% feels better in front of AutoML, as it feels more qualitative, less quantitative
\subsection{Interpretability}
As there is no clear definition for interpretability, we will consider it as ``the ability to provide \textit{explanations} in \textit{understandable terms} to a human'',
where \textit{explanations} are logical decision rules and \textit{understandable terms} relate to commonly used terms in the domain of the problem,
as suggested by \citet[chap. 1]{survey_NN_interpretability}.
Further, we use the term ``explanability'' in an exchangeable manner with ``interpretability'', as is commonly done.

(Why Interpretability is desirable?)
- in many ways desirable, as per \citet[pp. 3-4]{Zach2019InterpretabilityOD}, e.g.
  * \textbf{gaining trust}: most notable use case, e.g. medical diagnosis: predictions need to be validated by doctors \citep[p. 1]{review_NN_interpretability}
  * \textbf{model debugging}
    - model is optimised w.r.t. loss and judged on accuracy
    - might have unexpected performance drops in certain situations, makes model unreliable \citep[1B]{survey_NN_interpretability}
    - having an interpretable model + model induction process can help find ways to improve accuracy + reliability
  * \textbf{scientific understanding}: when only models can make sense of increasingly complex data anymore, interpretability enables extraction of learnt knowledge encoded
    in the model to make it accessible + reliable to humans (also mentioned here \citep[p. 1]{review_NN_interpretability})
  * \textbf{subconcious bias}, e.g. loan approval: must ensure that decision is not discriminatory
  * \textbf{regulatory}
    - such as EU's "Right to Explanation" warranted by the GDPR \citep[p. 1]{review_NN_interpretability}, \citep[1B]{survey_NN_interpretability} or
    - drug approval processes in situations where a machine learning model discovered a new drug, process needs to be transparent for the regulator to approve \citep[1B]{survey_NN_interpretability}

Commonly, methods for model interpretation are divided into intrinsic methods, where the search space only comprises models with a structure simple enough to be
considered ``explainable'' (such as tree-based or simple linear models), and post-hoc methods, where interpretation methods are applied after model training.
Amongst post-hoc techniques, we can further divide the space into model-specific (such as analysing GLM coefficients) and model-agnostic
(e.g. partial depence plots, ALE) methods. \citet[chap. 3.2]{molnar2022}

(Taxonomy)
\citet[chap. 2]{survey_NN_interpretability} extend this distinction to a 3-dimensional taxonomy, allowing for better categorisation of neural networks (NN), a model class
that in its fully-connected feedforwad form is inherently non-interpretable.
% alternative formulation
%As fully-connected feedforward neural networks (NNs) are inherently non-interpretable models, the taxonomy introduced by \citet[chap. 2]{survey_NN_interpretability} proves
%particulary useful to categorise NN explanability methods along 3 dimensions:
- \textbf{Passive vs Active Approaches}
  * passive: post-hoc
  * active: changing architecture or training process to make model interpretable
- \textbf{Type of Explanations}
  * examples: providing examples of what leads to desired output
  * attribution: attributing effect on output for a specific feature
  * hidden semantics: examine what type of inputs specific neurons / layers pick up on
  * rules: logical rules, e.g. if-then, trees
- \textbf{Local vs Global Interpretability}
  * local: explanation based on individual samples
  * semi-local: explanation based on set of samples, e.g. grouped together by some criterion
  * global: explaining network as a whole

(Evaluation)
- objective evaluation difficult, as no clear definition of interpretability
- \citet[3]{DoshiVelez2017TowardsAR} propose a taxonomy to categorise possible evaluation methods
  * \textbf{application-grounded evaluation}
    - most rigorous method, also most expensive + time consuming
    - idea: evaluate ``interpretability model'' directly w.r.t. the task -> human experts evaluate the outcome
    - e.g.: model performing medical diagnosis on patients would be evaluated by based on doctors doing the same
  * \textbf{human-grounded metrics}
    - similar to application-grounded, but tries to simplify task (while preserving its essence) so that laypersons can do it
    - cheaper due to larger, less qualified subject pool
    - especially suitable if only general concepts of the tasks need to be validated
    - examples for evaluation set-up
      * binary forced choice: human evaluator chooses which of two explanations he finds better
      * forward simulation: evaluator must correctly simulate model output when presented model input + explanation % TODO: perhaps binary forced choice is already enough to include, this (and the third one is a bit awkward to explain why this is good for interpretability imo)
  * \textbf{functionally-grounded evaluation}
    - assess explanation quality via some formally defined proxy for interpretability
    - no human time nor cost required beyond initial formulation of proxy
    - e.g. if we already have an interpretable model class (e.g. identified via human-grounded evaluation), can then rank different models of that class based on proxy
    - main challenge: what proxies

@@

(evolutionary algorithms)
- another model-free class of optimisation algorithms, inspired by naturally occuring evolution
- based on a population, which iteratively (iteration = generation) generates $\lambda$ offspring via 3 operators (all sub-bullets from \citet[pp. 10-]{genetic_algos})
  * \textbf{reproduction}: pick an individual to reproduce with p proportional to its fitness
  * \textbf{crossover}
    - select 2 ``parents'' from pool of reproducing individuals
    - for a position $k$ in their HP configs, with certain p, swap their values after $k$, yielding 2 children
    % TODO: use example from book page 12
  * \textbf{mutation}: with certain p, change values of an individual, e.g. by adding Gaussian noise (Gaussian mutation) to real values, or flipping bit for binary values
- selection at end of each genertion: keep $\mu$ best individuals (based on some fitness function), either from only offspring (``$(\mu,\lambda)$-selection'')
  or (more commonly) from population + offspring (``$(\mu+\lambda)$-selection'', guaranteed to keep best individual)
- pro: conceptually simple + can handle even complex parameter spaces (continuous, discrete, hierarchical, etc.) given appropriate implementation of operators
all above starting at HPO from \citep[chap. 1.3]{feurer_hyperparameter_2019}
- prominent examples
  * CMA-ES
    - ``Covariance Matrix Adapation Evolution Strategy''
    - offspring generation exclusively via multivariate normal \citep[p. 8]{hansen2023cmaevolutionstrategytutorial}
      * mean = weighted average of previous generation
      * covariance = weighted covariance of previous generation, with weighting as for mean
    - weighing scheme done to sample in a way as to reproduce previously successful (i.e. selected) steps \citep[p. 11]{hansen2023cmaevolutionstrategytutorial}
  * differential evolution \citep[-]{differential_evolution}
    - init population randomly so that entire param space is covered
      * $D$-dim vectors $\boldsymbol{x}_{i,G}$ with $i=1,2,...,n$ for gen $G$
    - mutation
      * for each $\boldsymbol{x}_{i,G}$, generate \textbf{mutant vector} $\boldsymbol{v}_{i,G+1}=\boldsymbol{x}_{r_1,G}+F\cdot(\boldsymbol{x}_{r_2,G}-\boldsymbol{x}_{r_3,G})$
      * $r_1,r_2,r_3\in\{1,2,...,n\}$ mutually different random idx + different from $i$
      * constant $F\in[0,2]$
    - crossover
      * \textbf{trial vector} $\boldsymbol{u}_{i,G+1}=(u_{i,G+1}^{(1)},u_{i,G+1}^{(2)},...,u_{i,G+1}^{(D)})$
        - chose random index $R\in\{1,2,...,D\}$
        - sample $p\sim U[0,1]$
        - define crossover constant $\text{CR}\in[0,1]$
        - $u_{i,G+1} = \begin{dcases}
          v_{i,G+1}^{(j)} & \text{if } p\le\text{CR}\text{ or } j=R \\
          x_{i,G}^{(j)} & \text{else}
         \end{dcases}$
    - selection: greedy, compare fitness of $\boldsymbol{u}_{i,G+1}$ with $\boldsymbol{x}_{i,G}$, pick better

@@

\subsection{Neural Architecture Search (NAS)}
- approaches specifically for NN HPO due to flexible structure of NNs and thus very large search space (each layer can have different \# nodes, different activations, for CNN different pooling, etc. operations)
- we don't employ NAS for our extension as research focusses on the NLP and image domains, % TODO: find some more references
  whose datasets exhibit strong correlation amongst features (i.e. tokens or pixels, respectively), an effect that is much weaker for tabular data \citep[p. 7499]{Borisov_2024}
- at least want to give a little overview over most notable approaches
  * \textbf{cell search space}
    - modularise NN architecture into cells, NN as chain-structure of those
    - cell = basic building block, fixed component of an NN
      * e.g. linear layer with specific \# of neurons in feedforward
      * or convolutional / pooling / etc layer for CNN
    - then optimise sequential placement of the cells as HPO problem
      * e.g. via random search or BO
      * \citet[p. 3]{zoph2017neuralarchitecturesearchreinforcement} + \citet[pp. 2-4]{Zoph_2018_CVPR} use RNN to recursively optimise HPs of a cell given already determined
        previous cells (+ HPs of current cell), which is trained via RL with the different HPs the RNN predicts being the actions and performance on held-out data being reward
    - \citep[chap. 3.2]{elsken_neural_2019}
  * \textbf{one-shot model}
    - trains one overall ``fabric'' comprising all architectures of search space
    - visualisation as DAG, nodes are nodes, edges in-between are operations on a node
    - each path through DAG (from input to output node) represents one architecture of search space -> train entire DAG, then pick optimal path (architecture)
    - pro: very efficient training, individual architectures share operations along edges they share, training one-shot model trains all subsumed models (more expensive
      than training a single model but less expensive than trying out all configurations included in the fabric)
    - \citep[pp. 1-2, p.8]{saxena2017convolutionalneuralfabrics}
    % don't necessarily talk about co-adaptation problem

@@

- we say a point dominates another

@@

  reference point (usually worst point in objective space)

@@

    - translate all but one objective into constraint, then optimise remaining objective under subject to the constraints

@@

  * define priorisation of objectives

@@

- aside from usualy baselines grid and random seach there are multi-objective BO adaptations, mainly using either of two approaches

@@

    - or one overall acquisition function aggregating surrogates, e.g. EHI, maximises expected improvement of hypervolume \citep[p. 16]{10.1145/3610536} + \citep[pp. 8f]{EHI}

@@

    - ranking mechanism based on 2 parts

@@

        (3) if individuals left, repeat from 1, increment rank \#

@@

        - individuals without two neighbour in any dimension are assigned infinitely high distance value

@@

  * ranking used for selecting $\mu$ best individuals to keep for next generation + for reproduction: via binary tournament solution: select two random individuals,
    pick best w.r.t. nds + cd for offspring creation (mutation / crossover)

@@

%likely shouldn't be an entire section
\section{EAGGA}
- NSGA-II inspired evolutionary + genetic algorithm
- uses AUC-ROC as performance metric and 3 interpretation metrics
  * NF: rel. \# of features used in model
  * NI: rel. \# of pairwise feature interaction effects
  * NNM: rel. \# of non-monotone feature effects
- aims to find high performing models (AUC) with low NF, low NI, and NNM -> ensures better interpretability than simply optimising for performance
  but creates very large objective space $\check{\boldsymbol\Lambda}$ to be optimised over with
  tuples $(\boldsymbol\lambda, \boldsymbol{s}, \boldsymbol{I}_s, \boldsymbol{m}_{\boldsymbol{I}_{\boldsymbol{s}}})=\check{\boldsymbol\lambda}\in\check{\boldsymbol\Lambda}$
  with
  * $\boldsymbol\lambda\in\boldsymbol\Lambda$ ``regular'' model HP config
  * $\boldsymbol{s}\in\{0,1\}^p$ vector denoting feature usage
  * $\boldsymbol{I}_s\in\{0,1\}^{p\times p}$ matrix denoting pairwise interactions
  * $\boldsymbol{m}_{\boldsymbol{I}_{\boldsymbol{s}}}\in\{-1,0,1\}^p$ vector denoting monotonicity constraint of each feature (-1 decreasing, 0 none, 1 increasing)
- thus would require to solve $\argmin_{\check{\boldsymbol\lambda}\in\check{\boldsymbol\Lambda}} (GE(\mathcal{I}_{\check{\boldsymbol\lambda}},\mathcal{D}),NF(\hat{f}_{\mathcal{D},\check{\boldsymbol\lambda}}),NI(\hat{f}_{\mathcal{D},\check{\boldsymbol\lambda}}),NNM(\hat{f}_{\mathcal{D},\check{\boldsymbol\lambda}}))$
- \citep[pp. 540f]{EAGGA}
- introduces equivalence relation $R$ ``allowed to interact'' on features that significantly reduces search space for more efficient optimization
- thus: augmented search space $\tilde{\boldsymbol\Lambda}=\boldsymbol\Lambda\times\mathcal{G}$ comprising model HP space and group structure space
  * each group structure $\boldsymbol{G}\in\mathcal{G}$ consists of $g$ groups
    - $G_1$ set of excluded features
    - $\forall k=2,...,g:G_k=(E_k,M_{E_k})$ tuple
      * $E_k$ set of features allowed to interact with each other
      * $M_{E_k}$ monotonicity constraint of entire group $k$
- EA on model hyperparameters $\boldsymbol\Lambda$
- GGA on group structure space $\mathcal{G}$
- \citep[pp. 541f]{EAGGA}
- futher introduces special initialisation of group structures to increase sample-efficiency of EAGGA algorithm compared to random init
- \citep[pp. 542f]{EAGGA}

@@

  * EAGGA algorithm implemented just as described in the paper

@@

    - NN total layers $\in\{3, ..., 10\}$, i.e. hidden layers $\in\{1, ..., 8\}$, init from trunc geom with p=0.5
    - NN nodes per hidden layer (for each group) $\in\{3, ..., 20\}$, init from trunc geom with p=0.5

@@

      * but instead of fitting 10 trees + taking rel. \# of features used as p for trunc geom, we simply use 0.5 (sklearn dectree examination not straightforward)

@@

        (e.g. if former is = total \# of features and one feature is indep. from target) in these edge cases we only use the features with non-0 filter values

@@

      * same as for feat detect: use p=0.5 to sample \# of interactions used instead of 10 dec trees

@@

      * use default HPs (max depth 30, minsplie = 20) of mlr3 classification tree implementation, as this is the library the paper uses

@@

    - included features are passed to dataset, dataset outputs included features, multiplied by the features' individual monotonicity (-1 / 1)

@@

      * hidden layers use ReLU activation and dropout afterwards, then a shared output layer with concatenated sub-NNs' outputs as input and sigmoid activation with
        (sigmoid implicit in the loss function for better numerical stability, NN itself outputs "logits", which refers to pre-activation output in pytorch)
      * loss is binary cross entropy loss, optimizer is AdamW with default params
    - feature sparsity thus achieved by only training on included feature groups -> goes somewhat against of deep learning where model is supposed to decide itself,

@@

    - monotonicty constraint achieved by clipping weights to [0, infty) for restricted equivalence relations (monotonic decrease achieved via dataset object multiplying
      features with their individual signs), bias clipping not necessary (constant additive term)

@@

        - hence weight clipping necessary, bias clipping not because only constant term

@@

      * after that, use patience of 100: if current model's loss is > mean of last 100 epochs' losses

@@

    -> add graphic of entire dataset split

@@

      * those show that there is no uptick in losses on the early stopping portion (disjunct from training portion) of the set, hence max reasonable

@@

      * i.e. datasets with p < 30 likely use all features, which translates to trunc geom prob = 1

@@

    - this cannot easily be inverted due to monotonicity constraints (weights cannot simply be multiplied by -1 if constraint)

@@

    - NI, NNM consequently (bounded by NF, can never be more than max \# for respective \# included features) rather low low

@@

  * crit was likely triggered by short-term spike, thus could potentially be resolved by comparing average of last k loss against average of >patience< losses prior

@@

  * compute on val sett, i.e. had 5 folds per individual

@@

\section{Conclusion and Future Outlook}
(Conclusion)
- tabular data still difficult discipline for neural networks
- research suggests heavy regularisation to improve performance on tabular data
- EAGGA proved to be successful in making XGBoost more interpretable while keeping performance on par with unregularised XGBoost
- we extend EAGGA to neural networks to see if we can utilise the regularisation induced by it to make NNs both interpretable and performant on tabular data
- as consequence propose new architecture allowing to model equivalence relations of EAGGA
- found overall performance comparable to that of XGBoost fitted using EAGGA, which is a plus, but no outperformance

(Future Outlook)
- MO BO on group structure space possible?

>>>>

% feels better in front of AutoML, as it feels more qualitative, less quantitative
\subsection{Interpretability}
As there is no clear definition for interpretability, we will consider it as ``the ability to provide \textit{explanations} in \textit{understandable terms} to a human'',
where \textit{explanations} are logical decision rules and \textit{understandable terms} relate to commonly used terms in the domain of the problem,
as suggested by \citet[chap. 1]{survey_NN_interpretability}. Further, we use the term ``explainability'' in an exchangeable manner with ``interpretability'',
as is commonly done.

% Why Interpretability is desirable?
Explainability of a model's reasoning is in many ways desirable. \citet[pp. 3-4]{Zach2019InterpretabilityOD} gives a range of examples,
amongst which are \textit{gaining trust}, e.g. when doctors rely on medical diagnosis predictions, avoiding \textit{subconcious biases} by
making sure loan approvals are non-discriminatory, or \textit{regulatory} reasons, most notably the EU's ``Right to Explanation'' warranted by the
GDPR \citep[p. 1]{review_NN_interpretability} or for approval of drugs discovered using machine learning models \citep[1B]{survey_NN_interpretability}.
It can further prove helpful for explaining unexpected drops in model performance, which could arise from optimising a model with respect to loss and then judging its
performance on a different metric such as accuracy, a practice known as \textit{model debugging} \citep[1B]{survey_NN_interpretability} or for
\textit{scientific understanding} in domains where only models can make sense of increasingly complex data anymore and learnt knowledge encoded in a model needs to be
made accessible to humans to be used reliably \citep[p. 1]{review_NN_interpretability}.

Commonly, methods for model interpretation are divided into intrinsic methods, where the search space only comprises models with a structure simple enough to be
considered ``explainable'' (such as tree-based or simple linear models), and post-hoc methods, where interpretation methods are applied after model training.
Amongst post-hoc techniques, we can further divide the space into model-specific (such as analysing GLM coefficients) and model-agnostic
(e.g. partial depence plots, ALE) methods \citep[chap. 3.2]{molnar2022}.

% Taxonomy
\citet[chap. 2]{survey_NN_interpretability} extend this distinction to a three-dimensional taxonomy, allowing for better categorisation of neural networks (NNs),
a model class that in its fully-connected feedforwad form is inherently non-interpretable.
- \textbf{Passive vs Active Approaches}
  * passive: post-hoc
  * active: changing architecture or training process to make model interpretable
- \textbf{Type of Explanations}
  * examples: providing examples of what leads to desired output
  * attribution: attributing effect on output for a specific feature
  * hidden semantics: examine what type of inputs specific neurons / layers pick up on
  * rules: logical rules, e.g. if-then, trees
- \textbf{Local vs Global Interpretability}
  * local: explanation based on individual samples
  * semi-local: explanation based on set of samples, e.g. grouped together by some criterion
  * global: explaining network as a whole

(Evaluation)
- objective evaluation difficult, as no clear definition of interpretability
- \citet[3]{DoshiVelez2017TowardsAR} propose a taxonomy to categorise possible evaluation methods
  * \textbf{application-grounded evaluation}
    - most rigorous method, also most expensive + time consuming
    - idea: evaluate ``interpretability model'' directly w.r.t. the task -> human experts evaluate the outcome
    - e.g.: model performing medical diagnosis on patients would be evaluated based on doctors doing the same
  * \textbf{human-grounded metrics}
    - similar to application-grounded, but tries to simplify task (while preserving its essence) so that laypersons can do it
    - cheaper due to larger, less qualified subject pool
    - especially suitable if only general concepts of the tasks need to be validated
    - examples for evaluation set-up
      * binary forced choice: human evaluator chooses which of two explanations he finds better
      * forward simulation: evaluator must correctly simulate model output when presented model input + explanation % TODO: perhaps binary forced choice is already enough to include, this (and the third one is a bit awkward to explain why this is good for interpretability imo)
  * \textbf{functionally-grounded evaluation}
    - assess explanation quality via some formally defined proxy for interpretability
    - no human time nor cost required beyond initial formulation of proxy
    - e.g. if we already have an interpretable model class (e.g. identified via human-grounded evaluation), can then rank different models of that class based on proxy
    - main challenge: what proxies

@@

(evolutionary algorithms)
- another model-free class of optimisation algorithms, inspired by naturally occuring evolution
- based on a population, which iteratively (iteration = generation) generates $\lambda$ offspring via 3 operators (all sub-bullets from \citet[pp. 10-]{genetic_algos})
  * \textbf{reproduction}: pick an individual to reproduce with prob proportional to its fitness
  * \textbf{crossover}
    - select 2 ``parents'' from pool of reproducing individuals
    - for a position $k$ in their HP configs, with certain prob, swap their values after position $k$, yielding 2 new children
    % TODO: use example from book page 12
  * \textbf{mutation}: with certain prob, change values of an individual, e.g. by adding Gaussian noise (Gaussian mutation) to real values, or flipping bit for binary values
- selection at end of each genertion: keep $\mu$ best individuals (based on some fitness function), either from only offspring (``$(\mu,\lambda)$-selection'')
  or (more commonly) from population + offspring (``$(\mu+\lambda)$-selection'', guaranteed to keep best individual)
- pro: conceptually simple + can handle even complex parameter spaces (continuous, discrete, hierarchical, etc.) given appropriate implementation of operators
all above starting at HPO from \citep[chap. 1.3]{feurer_hyperparameter_2019}
- prominent examples
  * CMA-ES
    - ``Covariance Matrix Adapation Evolution Strategy''
    - offspring generation exclusively via multivariate normal \citep[p. 8]{hansen2023cmaevolutionstrategytutorial}
      * mean = weighted average of previous generation
      * covariance = weighted covariance of previous generation, with weighting as for mean
    - weighing scheme done to sample in a way as to reproduce previously successful (i.e. selected) steps \citep[p. 11]{hansen2023cmaevolutionstrategytutorial}
  * differential evolution \citep[-]{differential_evolution}
    - init population randomly so that entire param space is covered
      * $D$-dim vectors $\boldsymbol{x}_{i,G}$ with $i=1,2,...,n$ for gen $G$
    - mutation
      * for each $\boldsymbol{x}_{i,G}$, generate \textbf{mutant vector} $\boldsymbol{v}_{i,G+1}=\boldsymbol{x}_{r_1,G}+F\cdot(\boldsymbol{x}_{r_2,G}-\boldsymbol{x}_{r_3,G})$
      * $r_1,r_2,r_3\in\{1,2,...,n\}$ mutually different random idx + different from $i$
      * constant $F\in[0,2]$
    - crossover
      * \textbf{trial vector} $\boldsymbol{u}_{i,G+1}=(u_{i,G+1}^{(1)},u_{i,G+1}^{(2)},...,u_{i,G+1}^{(D)})$
        - chose random index $R\in\{1,2,...,D\}$
        - sample $p\sim U[0,1]$
        - define crossover constant $\text{CR}\in[0,1]$
        - $u_{i,G+1} = \begin{dcases}
          v_{i,G+1}^{(j)} & \text{if } p\le\text{CR}\text{ or } j=R \\
          x_{i,G}^{(j)} & \text{else}
         \end{dcases}$
    - selection: greedy, compare fitness of $\boldsymbol{u}_{i,G+1}$ with $\boldsymbol{x}_{i,G}$, pick better

@@

\subsection{Neural Architecture Search (NAS)}
- approaches specifically for NN HPO due to flexible structure of NNs and thus very large search space (each layer can have different \# nodes, different activations, for CNN different pooling, etc. operations)
- we don't employ NAS for our extension as research focusses on the NLP and image domains, % TODO: find some more references
  whose datasets exhibit strong correlation amongst features (i.e. tokens or pixels, respectively), an effect that is much weaker for tabular data \citep[p. 7499]{Borisov_2024}
- at least want to give a little outline on most notable approaches
  * \textbf{cell search space}
    - modularise NN architecture into cells, NN as chain-structure of those
    - cell = basic building block, fixed component of an NN
      * e.g. linear layer with specific \# of neurons in feedforward
      * or convolutional / pooling / etc layer for CNN
    - then optimise sequential placement of the cells as HPO problem
      * e.g. via random search or BO
      * \citet[p. 3]{zoph2017neuralarchitecturesearchreinforcement} + \citet[pp. 2-4]{Zoph_2018_CVPR} use RNN to recursively optimise HPs of a cell given already determined
        previous cells (+ HPs of current cell), which is trained via RL with the different HPs the RNN predicts being the actions and performance on held-out data being reward
    - \citep[chap. 3.2]{elsken_neural_2019}
  * \textbf{one-shot model}
    - trains one overall ``fabric'' comprising all architectures of search space
    - visualisation as DAG, nodes are layers, edges in-between are operations on a layer's values
    - each path through DAG (from input to output node) represents one architecture of search space -> train entire DAG, then pick optimal path (architecture)
    - pro: very efficient training, individual architectures share operations along edges they share, training one-shot model trains all subsumed models (more expensive
      than training a single model but less expensive than trying out all configurations included in the fabric)
    - \citep[pp. 1-2, p.8]{saxena2017convolutionalneuralfabrics}
    % don't necessarily talk about co-adaptation problem

@@

- we say a ``point dominates another''

@@

  reference point (usually worst point in objective space), e.g. 0 for accuracy (optimum is 1)

@@

    - translate all but one objective into constraint, then optimise remaining objective subject to the constraints

@@

  * define priority of objectives

@@

- aside from usual baselines grid and random seach there are multi-objective BO adaptations, mainly using either of two approaches

@@

    - or one overall acquisition function aggregating surrogates, e.g. \textit{EHI}, maximises expected improvement of hypervolume \citep[p. 16]{10.1145/3610536} + \citep[pp. 8f]{EHI}

@@

    - multi-objective ranking mechanism based on 2 parts

@@

        (3) if any individuals left, repeat from 1, increment rank \#

@@

        - individuals without two neighbours in a dimension are assigned infinitely high distance value

@@

  * ranking used for selecting $\mu$ best individuals to keep for next generation + for reproduction: via binary tournament selection: select two random individuals,
    pick best w.r.t. nds + cd for offspring creation (i.e. mutation / crossover)

@@

%likely shouldn't be an entire section
\section{EAGGA}
- NSGA-II inspired evolutionary + genetic algorithm
- uses AUC-ROC as performance metric and 3 interpretation metrics
  * NF: rel. \# of features used in model
  * NI: rel. \# of pairwise feature interaction effects
  * NNM: rel. \# of non-monotone feature effects
- aims to find high performing models (AUC) with low NF, NI, and NNM -> ensures better interpretability than simply optimising for performance
  but creates very large objective space $\check{\boldsymbol\Lambda}$ to be optimised over with
  tuples $(\boldsymbol\lambda, \boldsymbol{s}, \boldsymbol{I}_s, \boldsymbol{m}_{\boldsymbol{I}_{\boldsymbol{s}}})=\check{\boldsymbol\lambda}\in\check{\boldsymbol\Lambda}$
  with
  * $\boldsymbol\lambda\in\boldsymbol\Lambda$ ``regular'' model HP config
  * $\boldsymbol{s}\in\{0,1\}^p$ vector denoting feature usage
  * $\boldsymbol{I}_s\in\{0,1\}^{p\times p}$ matrix denoting pairwise interactions
  * $\boldsymbol{m}_{\boldsymbol{I}_{\boldsymbol{s}}}\in\{-1,0,1\}^p$ vector denoting monotonicity constraint of each feature (-1 decreasing, 0 none, 1 increasing)
- thus would require to solve $\argmin_{\check{\boldsymbol\lambda}\in\check{\boldsymbol\Lambda}} (GE(\mathcal{I}_{\check{\boldsymbol\lambda}},\mathcal{D}),NF(\hat{f}_{\mathcal{D},\check{\boldsymbol\lambda}}),NI(\hat{f}_{\mathcal{D},\check{\boldsymbol\lambda}}),NNM(\hat{f}_{\mathcal{D},\check{\boldsymbol\lambda}}))$
- \citep[pp. 540f]{EAGGA}
- introduces equivalence relation $R$ ``allowed to interact'' on features that significantly reduces search space for more efficient optimization
- thus: augmented search space $\tilde{\boldsymbol\Lambda}=\boldsymbol\Lambda\times\mathcal{G}$ comprising model HP space and group structure space
  * each group structure $\boldsymbol{G}\in\mathcal{G}$ consists of $g$ groups:
    - $G_1$ set of excluded features
    - $\forall k=2,...,g:G_k=(E_k,M_{E_k})$ tuple
      * $E_k$ set of features allowed to interact with each other
      * $M_{E_k}$ monotonicity constraint of entire group $k$
- EA on model hyperparameters $\boldsymbol\Lambda$
- GGA on group structure space $\mathcal{G}$ with adapted mutation and crossover operators
- \citep[pp. 541f]{EAGGA}
- futher introduces special initialisation of group structures to increase sample-efficiency of EAGGA algorithm compared to random init
- \citep[pp. 542f]{EAGGA}

@@

  * EAGGA algorithm implemented largely as described in the paper

@@

    - NN total layers $\in\{3, ..., 10\}$, i.e. hidden layers $\in\{1, ..., 8\}$, init from trunc geom with prob=0.5
    - NN nodes per hidden layer (for each group) $\in\{3, ..., 20\}$, init from trunc geom with prob=0.5

@@

      * but instead of fitting 10 trees + taking rel. \# of features used as prob for trunc geom, we simply use 0.5 (sklearn dectree examination not straightforward)

@@

        (e.g. if former is = total \# of features and one feature is indep. from target) in these edge cases we simply use all features with non-0 filter values

@@

      * same as for feat detect: use prob=0.5 to sample \# of interactions used instead of 10 dec trees

@@

      * use default HPs (max depth 30, minsplit = 20) of mlr3 classification tree implementation, as this is the library the paper uses

@@

    - included features are passed to dataset, dataset outputs included features, multiplied by the features' individual monotonicity (-1 / 1), so that group monotonicity
      can be encoded with only $\{0,1\}$, as proposed in paper

@@

      * hidden layers use ReLU activation and dropout afterwards, then a shared output layer with concatenated sub-NNs' outputs as input and sigmoid activation
      * optimizer is AdamW with default params, loss is binary cross entropy loss (sigmoid thus implicit in the loss function for better numerical stability,
        NN itself outputs "logits", which refers to pre-activation output in pytorch)
    - feature sparsity thus achieved by only training on included feature groups -> goes somewhat against notion of deep learning where model is supposed to decide itself,

@@

    - monotonicty constraint achieved by clipping weights to [0, infty) after each epoch for restricted equivalence relations (monotonic decrease achieved
      via dataset object multiplying features with their individual signs), bias clipping not necessary (constant additive term)

@@

        - hence weight clipping necessary, bias clipping not because only constant additive term

@@

      * after that, use patience of 100: if current model's loss is > mean of last 100 epochs' losses -> stop early

@@

    -> refere Figure \ref{fig-dataset-split}

@@

      * those show that there is no uptick in losses on the early stopping portion (disjunct from training portion) of the set, hence max is reasonable

@@

      * i.e. datasets with p < 30 might use all features, which translates to trunc geom prob = 1

@@

    - for us this cannot easily be inverted due to monotonicity constraints (weights cannot simply be multiplied by -1 if constraint)

@@

    - NI, NNM consequently (bounded by NF, can never be more than max \# for respective \# included features) rather low

@@

  * crit was likely triggered by short-term spike, thus could potentially be resolved by comparing average of last k losses against average of >patience< losses prior

@@

  * compute on val set, i.e. had 5 folds per individual

@@

\section{Conclusion and Future Outlook}
(Conclusion)
- tabular data still difficult discipline for neural networks
- research suggests heavy regularisation to improve performance on tabular data
- EAGGA proved to be successful in making XGBoost more interpretable while keeping performance on par with unregularised XGBoost
- we extend EAGGA to neural networks to see if we can utilise the regularisation induced by it to make NNs both interpretable and performant on tabular data
- as consequence propose new architecture allowing to model equivalence relations of EAGGA
- ... conforming to \citet[chap. 2]{survey_NN_interpretability} taxonomy it fits as ..., ..., ... %TODO
- found overall performance comparable to that of XGBoost fitted using EAGGA, which is a plus, but no outperformance

(Future Outlook)
- MO BO on group structure space possible, perhaps via restricted BO?
